---
title: "Midterm Project"
output: github_document
Author: Na Yun Cho
---

```{r packages loaded}
library(tidyverse)
library(readxl)
library(patchwork)
```

# Problem 1 

The raw dataset contains weight data of three dogs, Raisin, Simone, and Gagne, per date ranging from 8/12/2018 to 9/7/2020. Each dog has two weight entries- weight in pounds and in ounces. The raw dataset also contains notes and weight data of a standard object that was weighed to ensure accuracy over time. 

The goals of this report are to examine the distribution of weights for each dog and the progression of each dog’s weight over time.


### (1) Tidy raw dataset
```{r}
dogs_raw = 
  read_excel("./data/dogweights.xlsx", range = "B2:G207", col_types = c("date", "text", "text", "text", "text",
  "text")) %>%
  rename(notes = ...6) %>%
  janitor::clean_names()
```

### (2) Produce first tidy dataframe
```{r}
dogs_df1 = 
  dogs_raw %>%
  drop_na(date) %>%
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(raisin = recode(raisin, '1810' = '18 10', '1803' = '18 03', '7 15' = '17 15')) %>%
  separate(raisin, into = c("raisinlbs", "raisinoz"), sep = " ") %>%
  mutate(raisin = as.numeric(raisinlbs) + as.numeric(raisinoz)/16) %>%
  separate(simone, into = c("simonelbs", "simoneoz"), sep = " " ) %>%
  mutate(simone = as.numeric(simonelbs) + as.numeric(simoneoz)/16) %>%
  separate(gagne, into = c("gagnelbs", "gagneoz"), sep = " " ) %>%
  mutate(gagne = as.numeric(gagnelbs) + as.numeric(gagneoz)/16) %>%
  mutate(
    year = as.factor(year),
    month = as.factor(month),
    day = as.factor(day)) %>%
  select(-c(raisinlbs:notes))
```

One of the issues of the imported raw dataset was the incorrect date formatting. 
So, I used 'col_types' to fix the issue.
Additionally, the columns were named incorrectly in the imported raw dataset. 
Thus, I set the import range of the raw data excel sheet to exclude the first row, and then renamed one column, and used 'clean_names()'.
Then, I used 'mutate' and 'recode' to fix errors in three of Raisin's weight entries.
I then used 'separate' and 'mutate' to combine the two weight data into one weight data per dog.
Finally, I produced the first tidy dataframe by using 'select' to leave out unnecessary columns. 

### (3) Produce second tidy dataframe
```{r}
dogs_df2 = 
  dogs_raw %>%
  drop_na(date) %>%
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(
    year = as.factor(year),
    month = as.factor(month),
    day = as.factor(day)) %>%
  select(-c(raisin:std))
```


### (4) Export both dataframes 
```{r}
write_csv(dogs_df1, "./data/problem1_dataframe1.csv")
write_csv(dogs_df2, "./data/problem1_dataframe2.csv")
```


# Problem 2
### (1) Identify the number of unique dates in the dataset
```{r}
unique_days = 
  dogs_df1 %>%
  group_by(year, month, day) %>%
  summarize(n_obs = n()) %>%
  nrow()
```
There are `r unique_days` unique dates in the dataset.
Raisin's maximum weight was `r max(dogs_df1$raisin, na.rm = TRUE)` lbs and minimum weight was `r min(dogs_df1$raisin, na.rm = TRUE)` lbs. 
In addition, Simone's maximum and minimum weights were `r max(dogs_df1$simone, na.rm = TRUE)` lbs and `r min(dogs_df1$simone, na.rm = TRUE)` lbs, respectively. 
Lastly, Gagne's maximum and minimum weights were `r max(dogs_df1$gagne, na.rm = TRUE)` lbs and `r min(dogs_df1$gagne, na.rm = TRUE)` lbs, respectively. 

### (2) Make a table 
```{r}
dogs_table = 
  data.frame(t(colSums(!is.na(dogs_df1)))) %>%
  select(raisin, simone, gagne)
rownames(dogs_table) = c("observations")

dogs_weight_mean = 
  dogs_df1 %>%
  summarize(
    raisin = mean(raisin, na.rm = TRUE),
    simone = mean(simone, na.rm = TRUE),
    gagne = mean(gagne, na.rm = TRUE))
rownames(dogs_weight_mean) = c("mean")

dogs_weight_sd = 
  dogs_df1 %>%
  summarize(
    raisin = sd(raisin, na.rm = TRUE),
    simone = sd(simone, na.rm = TRUE),
    gagne = sd(gagne, na.rm = TRUE))
rownames(dogs_weight_sd) = c("standard deviation")

dogs_table = do.call("rbind", list(dogs_table, dogs_weight_mean, dogs_weight_sd)) %>%
  knitr::kable(digits = 0.1)

dogs_table
```

# Problem 3
### Create a two-panel plot 
```{r}
dogs_df3 = 
  dogs_raw %>%
  drop_na(date) %>%
  mutate(raisin = recode(raisin, '1810' = '18 10', '1803' = '18 03', '7 15' = '17 15')) %>%
  pivot_longer(
    raisin:gagne, 
    names_to = "dog_name",
    values_to = "weights") %>%
  relocate(dog_name) %>%
  separate(weights, into = c("lbs", "oz"), sep = " ") %>%
  mutate(weight_in_lbs = as.numeric(lbs) + as.numeric(oz)/16) %>%
  select(-c(notes, std, lbs, oz,))

#Plot of distribution of weights for each dog 
dogs_weight_dist = 
  dogs_df3 %>%
  select(-date) %>%
  ggplot(aes(x = dog_name, y = weight_in_lbs, fill = dog_name)) + 
  geom_violin(alpha = 0.5) +
  stat_summary() +
  labs(
    title = "Distribution of weights per dog",
    x = "dog name", 
    y = "weights(lbs)")

#Plot of each dog's weight over time
dogs_weight_time =
  dogs_df3 %>%
  ggplot(aes(x = date, y = weight_in_lbs, color = dog_name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Each dog's weight over time",
    x = "date",
    y = "weights(lbs)") +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
  
dogs_weight_dist + dogs_weight_time
```

The left panel shows that the minimum and maximum weights of Raisin are much higher than any of the weights of Gagne and Simone. Raisin's weights show a right-skewed distribution with a mean around 18 lbs. 
However, Gagne's weight distribution is left-skewed with a mean around 10 lbs. 
Similarly, Simone's weight distribution also shows a mean around 10 lbs, but seems closer to approaching a normal distribution and has less of a spread than Gagne's weight distribution. 
It is apparent that Gagne's minimum weight is lower than that of Simone while their maximum weights seem very similar.

The right panel shows that there was a dramatic decrease in Raisin's weights from August 2018 to January 2019. 
Then, Raisin's weights have stayed fairly consistent between January 2019 and March 2020 until they showed a more increasing trend from March to September 2020. 
Simone's weights have stayed more consistent from August 2018 to September 2020. Although the weights seemed to slightly decrease from August 2018 to January 2019, from then on the weights stayed fairly consistent. 
However, Gagne's weights showed a clear decreasing trend from August 2018 to September 2019. 
The decrease in weights was more steep from around April 2019 to September 2019, most likely related to Gagne's imminent death.

```{r}
devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)
wordcountaddin::text_stats("p8105_mtp_nc2894.Rmd")
```

