---
title: "Midterm Project"
output: github_document
Author: Na Yun Cho
---

```{r}
library(tidyverse)
library(readxl)
library(patchwork)
```

# Problem 1 
### (1) Tidy raw dataset
```{r}
dogs_raw = 
  read_excel("./data/dogweights.xlsx", range = "B2:G207", sheet = "Sheet1", col_types = c("date", "text", "text", "text", "text",    "text")) %>%
  rename(notes = ...6) %>%
  janitor::clean_names()
```

### (2) Produce the first tidy dataframe
```{r}
dogs_df1 = 
  dogs_raw %>%
  drop_na(date) %>%
  separate(date, into = c("year", "month", "day"), sep = '-') %>%
  pivot_longer(
    raisin:gagne, 
    names_to = "dog_name",
    values_to= "weights") %>%
  relocate(dog_name) %>%
  separate(weights, into = c("lbs", "oz"), sep = ' ') %>%
  mutate(weight_in_lbs = as.numeric(lbs) + as.numeric(oz)/16) %>%
  mutate(
    year = as.factor(year),
    month = as.factor(month),
    day = as.factor(day)) %>%
  select(-c(notes, std, lbs, oz,))
```


### (3) Produce the second tidy dataframe that contains only the notes and dates
```{r}
dogs_df2 = 
  dogs_raw %>%
  drop_na(date) %>%
  separate(date, into = c("year", "month", "day"), sep = '-') %>%
  mutate(
    year = as.factor(year),
    month = as.factor(month),
    day = as.factor(day)) %>%
  select(-c(raisin:std))
```


# Problem 2
### (1) Identify the number of unique dates in the dataset
```{r}
unique_days_df = length(unique(dogs_raw$date))
```
There are `r unique_days_df` unique dates in the tidy dataset containing weights.


### (2) Make a well-formatted table that shows the number of observations per dog, and their average weight and standard deviation
```{r}
obsverations_per_dog_df = 
  dogs_df1 %>%
  group_by(dog_name) %>%
  summarize(
    n_obs = n(),
    mean_weights = mean(weight_in_lbs, na.rm = TRUE),
    sd_weights = sd(weight_in_lbs,na.rm = TRUE))
```

# Problem 3
### Create a two-panel plot 
```{r}
#Plot of distribution of weights for each dog 
weight_dist_per_dog = 
  dogs_df1 %>%
  select(-c(year:day)) %>%
  ggplot(aes(x = weight_in_lbs, y = dog_name)) + 
  geom_boxplot() +
  labs(
    title = "Distribution of weights for each dog",
    x = "Weight(lbs)", 
    y = "Dog name")

#Plot of each dog's weight over time
weight_time_per_dog =
  dogs_raw %>%
  drop_na(date) %>%
  pivot_longer(
    raisin:gagne, 
    names_to = "dog_name",
    values_to= "weights") %>%
  relocate(dog_name) %>%
  separate(weights, into = c("lbs", "oz"), sep = ' ') %>%
  mutate(weight_in_lbs = as.numeric(lbs) + as.numeric(oz)/16) %>%
  select(-c(std, notes, lbs, oz)) %>%
  ggplot(aes(x = date, y = weight_in_lbs, color = dog_name)) +
  geom_point(alpha = 0.5)+
  labs(
    title = "Each dog's weight over time",
    x = "Time",
    y = "Weights(lbs)")
  
  
  
```

